The \project{unRAVE} catalog of X.X million detailed stellar abundances for chemodynamic studies

### 2016-08-16 article/unrave.tex
Wrap wrap wrap

\begin{abstract}
On 14 September 2016 the \project{Gaia} mission will release positions, proper 
motions, and parallaxes for about two million stars.  Most studies intending to
make use of those data would benefit enormously by coupling that astrometry with
stellar chemical abundances.  The RAdial Velocity Experiment (\project{RAVE}) 
survey has acquired spectra for 292,036 stars in common with the 
first \project{Gaia} data release, constituting the largest overlap of any 
stellar spectroscopic sample.  Here we perform an independent analysis of the 
\project{RAVE} spectra using a new non-parametric implementation of \thecannon\ 
that employs prior probabilities on atomic line formation and stellar 
astrophysical parameters.  We deliver effective temperature $\teff$, surface 
gravity $\logg$, and chemical abundances of up to seven elements (Mg, Al, Si, 
Ca, Ti, Fe, Ni) for \Nstars\ stars.  We validate our results internally from 
repeat visits, and externally against existing high-resolution spectroscopic 
surveys.  The typical precision in chemical abundances is X.XX~dex.  
We provide improved distances by isochrone-fitting for \project{RAVE} stars
where astrometric solutions are not available in the first \project{Gaia}
data release.
\end{abstract}

@@

The Milky Way is considered to be the best laboratory for understanding galaxy formation and evolution.  This premise depends on the ability to precisely measure the position, motion, and chemistry of (many) individual stars, and to infer the structure, kinematics, and chemical enrichment of the Galaxy from those data \citep[e.g.,][]{Deason_2011,Ness_2012,Ness_2013a,Ness_2013b,Casey_2012,Casey_2013,Casey_2014a,Casey_2014b}.  % include references on: disk chemistry, halo structures
However, these quantities are not known for even 1\% of stars in the Milky Way.  Stellar distances are famously imprecise \citep[e.g.,][]{van_Leeuwen_2007,Jofre_2015,Madler_2016}, proper motions are frequently plagued by unquantified systematics from the first epoch observations \citep[e.g.,][]{Casey_Schlaufman_2015}, and stellar spectroscopists frequently report significantly different chemical abundance patterns \citep{Smiljanic_2014}.  If these measurements were known with high fidelity then \stub{xyz}.

>>>>

\begin{abstract}
On 14 September 2016 the \project{Gaia} mission will release positions, proper 
motions, and parallaxes for about two million stars.  Most studies intending to
make use of those data would benefit enormously by coupling that astrometry with
stellar chemical abundances.  The RAdial Velocity Experiment (\project{RAVE}) 
survey has acquired spectra for 292,036 stars in common with the 
first \project{Gaia} data release, constituting the largest overlap of any 
stellar spectroscopic sample.  Here we perform an independent analysis of the 
\project{RAVE} spectra using an implementation of \thecannon\ 
that uses prior probabilities on atomic line formation and stellar 
astrophysical parameters.  We deliver effective temperature $\teff$, surface 
gravity $\logg$, and chemical abundances of up to seven elements (Mg, Al, Si, 
Ca, Ti, Fe, Ni) for \Nstars\ stars.  We validate our results internally from 
repeat visits, and externally against existing high-resolution spectroscopic 
surveys.  The typical precision in chemical abundances is X.XX~dex.  
We provide improved distances by isochrone-fitting for \project{RAVE} stars
where astrometric solutions are not available in the first \project{Gaia}
data release.
\end{abstract}

@@

The Milky Way is considered to be the best laboratory for understanding galaxy
formation and evolution.  This premise hinges on the ability to precisely measure 
the position, motion, and chemistry of (many) individual stars, and to infer the 
structure, kinematics, and chemical enrichment of the Galaxy from those data 
\citep[e.g.,][]{Deason_2011,Ness_2012,Ness_2013a,Ness_2013b,Casey_2012,Casey_2013,
Casey_2014a,Casey_2014b}.  % include references on: disk chemistry, halo structures
However, these quantities are not known for even 1\% of stars in the Milky Way.  
Stellar distances are famously imprecise \citep[e.g.,][]{van_Leeuwen_2007,
Jofre_2015,Madler_2016}, proper motions can be plagued by unquantified systematics
from the first epoch observations \citep[e.g.,][]{Casey_Schlaufman_2015}, and 
stellar spectroscopists frequently report significantly different chemical 
abundance patterns \citep{Smiljanic_2014}.  Countless long-standing scientific
questions could be definitively answered if these measurements were known with
high fidelity: \stub{x? \citep{someone}; y; and z, to list just a few}.
